---
title: "Final"
author: "Kjunwonl"
date: "12/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(moderndive)
library(ggplot2)
library(dplyr)
library(data.table)
```

## R Markdown

```{r}
lifting <- fread("openpowerlifting-2019-11-20.csv", sep="auto")
```
#adding binary result of 0 = no and 1 = yes for equipments and testing columns, also F = 1 M = 0
```{r}
set_equip <- function(str){
  typeof(str)
  #for (i in length(str)) {
    if(str == "Wraps") return (3)
    else if (str == "Raw") return (0)
    else if (str == "Single-ply") return(1)
    else return (2)
  #}
}

lifting <-  lifting [, Equip := mapply(set_equip,Equipment)]


lifting$test <- as.numeric(lifting$Tested == "Yes")
lifting$Gender <- as.numeric(lifting$Sex == "F")
```

#Get rid of irrelevant columns like Squat4Kg, Bench4Kg, Deadlift4Kg, Tested (categorical not the binary one I created), BirthYearClass and Country
```{r}
lifting[,c("Squat4Kg", "Bench4Kg", "Deadlift4Kg", "Tested", "Country", "BirthYearClass"):=NULL]
```

#getting rid of empty rows with empty cells
```{r}
lifting[lifting==""] <- NA
lifting <- na.omit(lifting)
```

1.
#Cluster Analysis: for body weight
since all federations use different weight classes, I am going to select the most commonly used federation.

```{r}
sort(table(lifting[,28]))
```
USAPL most commonly used federation so I will filter my data to only that

```{r}
library(stringr)

lifting.x = lifting[,c(8, 9, 28)]
#remove all + in weightClassKg
lifting.x = lifting.x[,WeightClassKg := str_replace_all(WeightClassKg,"\\+","")]
#Use data only from the USAPL federation
lifting.x <- lifting.x[Federation == "USAPL"]
#convert to numeric for weight classes
lifting.x$WeightClassKg <- as.numeric(as.character(lifting.x$WeightClassKg))
#removing the Federation column for clustering method since it is categorical
lifting.x[,c("Federation"):=NULL]
lifting.x[lifting.x==""] <- NA
lifting.x <- na.omit(lifting.x)

kmeans.result = kmeans(lifting.x, centers=15) # we did no centering or scaling here. as usual, consider if those are needed for your data
kmeans.result$centers
kmeans.result$cluster
centers = kmeans.result$centers[kmeans.result$cluster, ] 
distances = sqrt(rowSums((lifting.x - centers)^2))
outliers = order(distances, decreasing=T)[1:20]

pch = rep(1, nrow(lifting.x))
pch[outliers] = 2:(length(outliers)+1)
col <- rep("red", nrow(lifting.x))
col[outliers] = "blue"
pairs(lifting.x, pch=pch, col=col)
```

```{r}
kmeans.result
print(outliers)
```

```{r}
ggplot(lifting.x, aes(x=BodyweightKg, y=WeightClassKg, color=col, pch=pch)) + geom_point() + scale_shape_identity()
```

2.
#Analysis: PCA
How does body weight, avg squat, avg bench, and avg deadlift relate to each other?
```{r}
factor_list <- c("Equip", "Best3SquatKg", "Best3BenchKg", "Best3DeadliftKg")
cols<- factor_list
lifting1 <- lifting

lifting2 <- lifting1[, (cols) := lapply(.SD, function(x) (x-mean(x))/sd(x) ),.SDcols = cols ]
```

```{r}
lifting2[,c("Equip", "Best3SquatKg", "Best3BenchKg", "Best3DeadliftKg")]
```

```{r}
lifting.pca <- prcomp(lifting2[,c("Equip", "Best3SquatKg", "Best3BenchKg", "Best3DeadliftKg")], center = TRUE,scale. = TRUE)


summary(lifting.pca)
```

```{r}
str(lifting.pca)
lifting.pca$sdev 
lifting.pca.var = lifting.pca$sdev^2
pve = 100 * (lifting.pca.var / sum(lifting.pca.var))
perc_data = data.table(percent=pve, PC=1:length(pve))
ggplot(perc_data, aes(x=PC, y=percent)) + 
  geom_line() +
  geom_text(aes(label=round(percent, 2)), size=4, vjust=-.5) +
  ylab("Proportion of Variance Explained\n") +
  xlab("Principle Component")
```
```{r}
lifting.pca$x #this data does not give us the categories, we can add them
pca_data <- data.frame(lifting.pca$x, Species=lifting$AgeClass)
head(pca_data)
```

```{r}
ggplot(pca_data, aes(x=PC1, y=PC2, color=Species)) + geom_point()
```
```{r}
lifting.pca$rotation
rotation_data <- data.table(lifting.pca$rotation, variable=row.names(lifting.pca$rotation))
ggplot(rotation_data) + 
  geom_segment(aes(xend=PC1, yend=PC2), x=0, y=0, arrow=arrow(length = unit(0.05, "inches"), type = "closed")) + 
  geom_text(aes(x=PC1, y=PC2, label=variable), hjust=0, size=3, color='red') + 
  xlim(-1.,1) + 
  ylim(-1.,1.)
```

```{r}
factor_list <- c("test", "Best3SquatKg", "Best3BenchKg", "Best3DeadliftKg")
cols<- factor_list
lifting3 <- lifting
lifting4 <- lifting3[, (cols) := lapply(.SD, function(x) (x-mean(x))/sd(x) ),.SDcols = cols ]
```

```{r}
lifting4[,c("test", "Best3SquatKg", "Best3BenchKg", "Best3DeadliftKg")]
```

```{r}
lifting.pca2 <- prcomp(lifting4[,c("test", "Best3SquatKg", "Best3BenchKg", "Best3DeadliftKg")], center = TRUE,scale. = TRUE)


summary(lifting.pca2)
```

```{r}
str(lifting.pca2)
lifting.pca2$sdev 
lifting.pca.var2 = lifting.pca2$sdev^2
pve2 = 100 * (lifting.pca.var2 / sum(lifting.pca.var2))
perc_data2 = data.table(percent=pve2, PC=1:length(pve2))
ggplot(perc_data2, aes(x=PC, y=percent)) + 
  geom_line() +
  geom_text(aes(label=round(percent, 2)), size=4, vjust=-.5) +
  ylab("Proportion of Variance Explained\n") +
  xlab("Principle Component")
```

```{r}
lifting.pca2$x #this data does not give us the categories, we can add them
pca_data2 <- data.frame(lifting.pca2$x, Species=lifting$AgeClass)
head(pca_data2)
```

```{r}
ggplot(pca_data2, aes(x=PC1, y=PC2, color=Species)) + geom_point()
```

```{r}
lifting.pca2$rotation
rotation_data2 <- data.table(lifting.pca2$rotation, variable=row.names(lifting.pca2$rotation))
ggplot(rotation_data2) + 
  geom_segment(aes(xend=PC1, yend=PC2), x=0, y=0, arrow=arrow(length = unit(0.05, "inches"), type = "closed")) + 
  geom_text(aes(x=PC1, y=PC2, label=variable), hjust=0, size=3, color='red') + 
  xlim(-1.,1) + 
  ylim(-1.,1.)
```
#Variable/Data Analysis:

```{r}
library(psych)

lifting <- fread("openpowerlifting-2019-11-20.csv", sep="auto")
set_equip <- function(str){
  typeof(str)
  #for (i in length(str)) {
    if(str == "Wraps") return (3)
    else if (str == "Raw") return (0)
    else if (str == "Single-ply") return(1)
    else return (2)
  #}
}

lifting <-  lifting [, Equip := mapply(set_equip,Equipment)]


lifting$test <- as.numeric(lifting$Tested == "Yes")
lifting$Gender <- as.numeric(lifting$Sex == "F")

lifting[,c("Squat4Kg", "Bench4Kg", "Deadlift4Kg", "Tested", "Country", "BirthYearClass"):=NULL]

lifting[lifting==""] <- NA
lifting <- na.omit(lifting)
```

```{r}
ggplot(lifting, aes(Age)) + geom_histogram(color = "white", binwidth = 4)
```
```{r}
summary(lifting$Age)
```

```{r}
ggplot(lifting, aes(x = AgeClass, y = Wilks)) + geom_col()
```

```{r}
ggplot(lifting, aes(AgeClass, TotalKg, fill = Equip)) + geom_bar(stat = "identity", position = "dodge")
```

```{r}
ggplot(lifting, aes(AgeClass, TotalKg, fill = test)) + geom_bar(stat = "identity", position = "dodge")
```
